### 2023-04-05 bachelorthesis/02-theory.tex
#39: Adjust chapter style, Seperate chapters in latex files, further Adjustment of brief commmunication
#39: Correction of brief communication

\chapter{Theory}
\label{chap:theory}

In the following the box size is increased relative to the standard box size $(L_\xcoord,~L_\ycoord) = (76.27,~89.76)\,\rhoth$ in the radial and binormal direction. Here, $\xcoord$ is the radial coordinate that labels the flux surfaces normalized by the thermal Larmor radius $\rhoth$, $\ycoord$ labels the field lines and is an approximate binormal or toroidal coordinate. Together with the coordinate $s$ which parameterizes the length along the field lines and represents the poloidal coordinate these quantities form the Hamada coordinates \cite{hamada1958}.
The increased box sizes are indicated by the real parameter $\NR$ for radial and $\NB$ for the binormal direction with the nomenclature $\NR\times \NB$ throughout this work.
Note that, the number of modes in the respective direction, i.e., $N_x$ and $N_m$, respectively, is always adapted accordingly to retain a spatial resolution compliant to the standard resolution [Tab.~\ref{tab:resolution}] and standard box size. \\
The $\exb$ staircase pattern is manifest as radial structure formation in the $\exb$ shearing rate defined by\cite{rath2016, doi:10.1063/1.3005380, rath2016, peeters2016}
\begin{equation}
	\wexb = \frac{1}{2} \frac{\partial^2 \langle \phi \rangle}{\partial \xcoord^2},
	\label{eq:shearingrate}
\end{equation}
where $\langle \phi \rangle$ is the zonal electrostatic potential normalized by $\rho_\ast T/e$ ($\rho_\ast = \rhoth/R$ is the thermal Larmor radius normalized with the major radius $R$, $T$ is the temperature, $e$ is the elementary charge).
The zonal potential is calculated from the electrostatic potential $\phi$ on the two-dimensional $\xcoord$-$\ycoord$-plane at the low field side according to\cite{rath2021}
\begin{equation}
\langle \phi \rangle = \frac{1}{L_\ycoord} \int_0^{L_\ycoord} \mathrm{d}\ycoord ~ \phi(\xcoord,\ycoord,s=0).
\end{equation}
The E$\times$B shearing rate $\wexb$ is the radial derivative of the advecting zonal flow velocity \cite{doi:10.1063/1.871313, doi:10.1063/1.872847} and quantifies the zonal flow induced shearing of turbulent structures \cite{doi:10.1063/1.859529, doi:10.1063/1.871313, doi:10.1063/1.872367}. \\
Consistent with Ref.~\cite{peeters2016} the turbulence level is quantified by the turbulent heat conduction coefficient $\chi$, which is normalized by $\rhoth^2 \vth/R$ ($\vth = \sqrt{2 T/m}$ is the thermal velocity and $m$ is the mass). Furthermore, quantities $\rhoth$, $R$, $T$, $\vth$ and $m$ are referenced quantities from Ref.\cite{peeters2016,peeters2009}.
\newpage
In order to diagnose the temporal evolution of the staircase pattern and to obtain an estimate of its amplitude the radial Fourier transform of the $\exb$ shearing rate is considered. 
It is defined by
\begin{equation}
	\wexb = \sum_{\kzf} \hatwexb(\kzf,t) \, \exp(\mathrm{i} \kzf \xcoord),
	\label{eq:shearingrate_fourier}
\end{equation}
where $\hatwexb$ is the complex Fourier coefficient and \linebreak $\kzf = 2\pi \nzf/L_\xcoord$
% explicitly introduce the terms zonal flow wave vector and zonal flow mode number for later usage
defines the zonal flow wave vector with the zonal flow mode number $\nzf$ ranging in $-(N_\xcoord -1)/2 \leq \nzf \leq (N_\xcoord -1)/2 $.
% define the zonal flow modes' amplitude in terms of the shearing rate
Based on the definitions above, the shear carried by the zonal flow mode with wave vector $\kzf$ is defined by $\hatwexbamp = 2 |\hatwexb|$. 
% introduces the term basic mode of the pattern for later usage
In general, the zonal flow mode that dominates the $\exb$ staircase pattern, also referred to as the \textit{basic mode} of the pattern in this work, exhibits the maximum amplitude in the spectrum $\hatwexbamp$.\\

>>>>

\chapter{Theory}
\label{chap:theory}
\thispagestyle{empty}
\newpage

In the following the box size is increased relative to the standard box size $(L_\xcoord,~L_\ycoord) = (76.27,~89.76)\,\rhoth$ in the radial and binormal direction. Here, $\xcoord$ is the radial coordinate that labels the flux surfaces normalized by the thermal Larmor radius $\rhoth$, $\ycoord$ labels the field lines and is an approximate binormal coordinate. Together with the coordinate $s$ which parameterizes the length along the field lines and is referred to as the parallel coordinate these quantities form the Hamada coordinates \cite{hamada1958}.
The increased box sizes are indicated by the real parameter $\NR$ for radial and $\NB$ for the binormal direction with the nomenclature $\NR\times \NB$ throughout this work.
Note that, the number of modes in the respective direction, i.e., $N_x$ and $N_m$, respectively, is always adapted accordingly to retain a spatial resolution compliant to the standard resolution [Tab.~\ref{tab:resolution}] and standard box size. \\
The $\exb$ staircase pattern is manifest as radial structure formation in the $\exb$ shearing rate defined by\cite{rath2016, doi:10.1063/1.3005380, rath2016, peeters2016}
\begin{equation}
	\wexb = \frac{1}{2} \frac{\partial^2 \langle \phi \rangle}{\partial \xcoord^2},
	\label{eq:shearingrate}
\end{equation}
where $\langle \phi \rangle$ is the zonal electrostatic potential normalized by $\rho_\ast T/e$ ($\rho_\ast = \rhoth/R$ is the thermal Larmor radius normalized with the major radius $R$, $T$ is the temperature, $e$ is the elementary charge).
The zonal potential is calculated from the electrostatic potential $\phi$ on the two-dimensional $\xcoord$-$\ycoord$-plane at the low field side according to\cite{rath2021}
\begin{equation}
\langle \phi \rangle = \frac{1}{L_\ycoord} \int_0^{L_\ycoord} \mathrm{d}\ycoord ~ \phi(\xcoord,\ycoord,s=0).
\end{equation}
The E$\times$B shearing rate $\wexb$ is the radial derivative of the advecting zonal flow velocity \cite{doi:10.1063/1.871313, doi:10.1063/1.872847} and quantifies the zonal flow induced shearing of turbulent structures \cite{doi:10.1063/1.859529, doi:10.1063/1.871313, doi:10.1063/1.872367}. \\
Consistent with Ref.~\cite{peeters2016} the turbulence level is quantified by the turbulent heat conduction coefficient $\chi$, which is normalized by $\rhoth^2 \vth/R$ ($\vth = \sqrt{2 T/m}$ is the thermal velocity and $m$ is the mass). Furthermore, quantities $\rhoth$, $R$, $T$, $\vth$ and $m$ are referenced quantities from Ref.\cite{peeters2016,peeters2009}.
\newpage
In order to diagnose the temporal evolution of the staircase pattern and to obtain an estimate of its amplitude the radial Fourier transform of the $\exb$ shearing rate is considered. 
It is defined by
\begin{equation}
	\wexb = \sum_{\kzf} \hatwexb(\kzf,t) \, \exp(\mathrm{i} \kzf \xcoord),
	\label{eq:shearingrate_fourier}
\end{equation}
where $\hatwexb$ is the complex Fourier coefficient and \linebreak $\kzf = 2\pi \nzf/L_\xcoord$
% explicitly introduce the terms zonal flow wave vector and zonal flow mode number for later usage
defines the zonal flow wave vector with the zonal flow mode number $\nzf$ ranging in $-(N_\xcoord -1)/2 \leq \nzf \leq (N_\xcoord -1)/2 $.
% define the zonal flow modes' amplitude in terms of the shearing rate
Based on the definitions above, the shear carried by the zonal flow mode with wave vector $\kzf$ is defined by $\hatwexbamp = 2 |\hatwexb(\kzf,t)|$. 
% introduces the term basic mode of the pattern for later usage
In general, the zonal flow mode that dominates the $\exb$ staircase pattern, also referred to as the \textit{basic mode} of the pattern in this work, exhibits the maximum amplitude in the spectrum $\hatwexbamp$.\\

### 2023-04-05 bachelorthesis/04-results.tex
#39: Correction of brief communication

% what kind of analysis
In a third test the binormal box size is varied with the radial box size fixed to $\NR = 3$.
This test covers the realizations:
\begin{gather*}
	\boxed{\NR \times \NB \in [3\times1.5,~3\times2.5,~3\times3,~3\times5]}~.
\end{gather*}
As in the isotropic scan the turbulence subdued and a fully developed staircase pattern formats after $\sim 2000\,R/\vth$ [Fig.~\ref{fig:eflux-3x1.5-2.5-3-5-comparison}]. The convergence of staircase pattern can be seen in Fig.~\ref{fig:wexb-stable-comparison}\,(c) and confirms again a size of a typical mesoscale. Fig.~\ref{fig:wexb-stable-comparison}\,(c) also confirms that indeed a competition between patterns with two sizes $\lambda \in [57.20,~ 76.27]\,\rhoth$ causing the different results for $3 \times 1$ and $3\times 3$. The zonal flow mode number varies between $\nzf = 3,4$ which can be seen in Fig.~\ref{fig:wexb-stable-comparison}\,(c) in the $3\times 2.5$ realization. The staircase structure has a pattern between $3$ and $4$ repetitions which get represented in the second repetition with no positive shear. Instead the pattern returns immediately after reaching the maximum shear ($+ \gamma$) to the minimum shear ($- \gamma$) of the third repetition in a steep flank. The Fourier analysis of this case yields no definitely basic mode rather two dominate modes with $\nzf = 3, 4$ with a fraction of the maximum amplitude $\hatwexbamp$ each (not shown).
\includegraphicsHere{{Comparison/Boxsize/S6_rlt6.0_boxsize3x1-1.5-2.5-3-5_Ns16_Nvpar48_Nmu9_comparison}.pdf}{
	Time traces of the heat conduction coefficient $\chi$ for $\rlt = 6.0$ for binormal increased box sizes
}{fig:eflux-3x1.5-2.5-3-5-comparison}{}\bigskip

@@

In the final test the inverse background temperature gradient length $\rlt$ is varied at fixed $3\times3$ box size.
Since suppression of turbulence usually occurs at later times when approaching the finite heat flux threshold from below \cite{peeters2016}, the analysis aims to lengthen the phase during which the zonal flow varies in time due to turbulent Reynolds stresses.
This scan covers realizations with:
\begin{gather*}
	\boxed{\rlt \in [6.0,~6.2,~6.4]}~.
\end{gather*}
% what is the finite heat flux threshold
In the case of $\rlt = 6.2$ turbulence suppression is observed for $t > 12000\,R/\vth$, while stationary turbulence during the entire simulation time trace of $12000\,R/\vth$ is found for $\rlt = 6.4$.
The finite heat flux threshold, hence, is:
\begin{gather*}
	\boxed{\rlt|_\mathrm{finite} = 6.3 \pm 0.1}~.
\end{gather*}
% Which stationay radial size occurs????
Although the initial quasi-stationary turbulence in the former case is significantly larger compared to the $\rlt = 6.2$ realization discussed in the second test, a stationary pattern with basic zonal flow mode $\nzf = 3$ establishes. 
% Does it confirm our previous tests
Again, the $\nzf = 1$ (box scale) zonal flow mode does not grow secularly during the entire turbulent phase.
Also, this test confirms the statistical soundness of the converged pattern size of $\sim 57.20 - 76.27\,\rhoth$.\\

>>>>

% what kind of analysis
In a third test the binormal box size is varied with the radial box size fixed to $\NR = 3$.
This test covers the realizations:
\begin{gather*}
	\boxed{\NR \times \NB \in [3\times1.5,~3\times2.5,~3\times3,~3\times5]}~.
\end{gather*}
As in the isotropic scan the turbulence subdued and a fully developed staircase pattern forms after $\sim 2000\,R/\vth$ [Fig.~\ref{fig:eflux-3x1.5-2.5-3-5-comparison}]. The convergence of staircase pattern can be seen in Fig.~\ref{fig:wexb-stable-comparison}\,(c) and confirms again a size of a typical mesoscale. Fig.~\ref{fig:wexb-stable-comparison}\,(c) also confirms that indeed a competition between patterns with two sizes $\lambda \in [57.20,~ 76.27]\,\rhoth$ causing the different results for $3 \times 1$ and $3\times 3$. The zonal flow mode number varies between $\nzf = 3,4$ which can be seen in Fig.~\ref{fig:wexb-stable-comparison}\,(c) in the $3\times 2.5$ realization. The staircase structure has a pattern between $3$ and $4$ repetitions which get represented in the second repetition with no signifciant plateau at positive shear. Instead the pattern returns immediately after reaching the maximum shear ($+ \gamma$) to the minimum shear ($- \gamma$) of the third repetition in a steep flank. The Fourier analysis of this case yields no definitely basic mode rather two dominating modes with $\nzf = 3, 4$ with a fraction of the maximum amplitude $\hatwexbamp$ each (not shown).
\includegraphicsHere{{Comparison/Boxsize/S6_rlt6.0_boxsize3x1-1.5-2.5-3-5_Ns16_Nvpar48_Nmu9_comparison}.pdf}{
	Time traces of the heat conduction coefficient $\chi$ for $\rlt = 6.0$ for binormal increased box sizes
}{fig:eflux-3x1.5-2.5-3-5-comparison}{}\bigskip

@@

In the final test the inverse background temperature gradient length $\rlt$ is varied at fixed $3\times3$ box size.
Since suppression of turbulence usually occurs at later times when approaching the finite heat flux threshold from below \cite{peeters2016}, the analysis aims to lengthen the phase during which the zonal flow varies in time due to turbulent Reynolds stresses.
This scan covers realizations with:
\begin{gather*}
	\boxed{\rlt \in [6.0,~6.2,~6.4]}~.
\end{gather*}
% what is the finite heat flux threshold
In the case of $\rlt = 6.2$ turbulence suppression is observed for $t > 11000\,R/\vth$, while stationary turbulence during the entire simulation time trace of $12000\,R/\vth$ is found for $\rlt = 6.4$.
The finite heat flux threshold, hence, is:
\begin{gather*}
	\boxed{\rlt|_\mathrm{finite} = 6.3 \pm 0.1}
\end{gather*}
in accordance to Ref. \cite{peeters2016}.
% Which stationay radial size occurs????
Although the initial quasi-stationary turbulence in the former case is significantly longer compared to the $\rlt = 6.2$ realization discussed in the second test, a stationary pattern with basic zonal flow mode $\nzf = 3$ establishes. 
% Does it confirm our previous tests
Again, the $\nzf = 1$ (box scale) zonal flow mode does not grow secularly during the entire turbulent phase.
Also, this test confirms the statistical soundness of the converged pattern size of $\sim 57.20 - 76.27\,\rhoth$.\\
